---
title: "Human Activity Prediction Assignment"
author: "Sriram Chitturi"
date: "January 27, 2016"
output: html_document
---

##Introduction

This project uses the data from 6 participants who had accelerometers on the belt, forearm, arm, and dumbell while exercising. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Training data - https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

Test data - https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Data set is described in detail at this link - http://groupware.les.inf.puc-rio.br/har

The goal of the project is to predict the manner in which they did the exercise which is the "Classe" variable in the data set.

## Loading and cleaning the data

```{r message=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
```

```{r}
set.seed(12345); # set seed for reproducibility

# Load data
pmldata <- read.csv(file="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("NA", ""));
```
There are 160 columns. We can eliminate the columns which are not useful.
After looking at the summary of the dataset using __str(training)__ function I see a lot of NAs in some columns.
First eliminating these columns from the training set.
```{r}
# get the percentage of rows in each column with NAs
na_count <-sapply(pmldata, function(y) sum(length(which(is.na(y)))));
na_count <- na_count/nrow(pmldata)*100; # calculate percentage
columns_na <- names(na_count[ which(na_count > 50)]); # eliminate more than 90% NAs

#eliminating these columns from the data set
pmldata <- pmldata[, -which(names(pmldata) %in% columns_na)];

```


From the remaining columns eliminating some unnecessary columns like timestamps and usernames.
```{r}
pmldata <- pmldata[, -which(colnames(pmldata) %in% c("X", "user_name", "raw_timestamp_part_1",
                      "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window"))];
```

Set columns type to double.
```{r}
id <- c(1,3:ncol(pmldata)-1);
pmldata[,id] <- as.double(as.character(unlist(pmldata[,id])));
str(pmldata);
```

##Training and predicting

```{r message=FALSE}

# splitting the dataset into 60/40 
inTrain <- createDataPartition(pmldata$classe, p=0.60, list=F);
training <- pmldata[inTrain,];
testing <- pmldata[-inTrain,];
dim(training);
dim(testing);
```

```{r}
modelFit <- train(classe~., data=training, method="rf", trControl = trainControl(number=2, repeats = 2))
plot(modelFit);
```

##Verification of model
```{r}
predictions <- predict(modelFit, newdata=testing);
confusionMatrix(predictions, testing$classe);
```
We see that the model predicted with accuracy of 99%.

##Final prediction on testing data for quiz.

```{r}
finalTesting <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
                         na.strings = c("NA", ""));

id <- c(1,3:ncol(finalTesting)-1);
finalTesting[,id] <- as.double(as.character(unlist(finalTesting[,id])));

predict(modelFit, newdata=finalTesting);

```
